Shallow EDSLs and Object-Oriented Programming

### 2016-06-21 paper/sections/Interpretations.tex
Shorten S4

\begin{comment}
Weixin writes this one.

Go over jeremy's examples, maybe having only 3 diagram 
constructs instead of 5 for space reasons.

Use five constructs to show extensibility.

Think of how to introduce our tool? Using the Jeremy's examples? 
or introducing before with some other examples?

\end{comment}

@@

Conversely, we can define introduce new interpretation in a
modular and intuitive way with an OO language like Scala:
\begin{figure}
\lstinputlisting[linerange=2-15]{../src/wellsized/Circuit.scala}%APPLY:MULTIPLE_SCALA
\caption{Adding new interpretations}
\label{code:operation}
\end{figure}
Instead of modifying the original code, we define \lstinline{wellSized} modularly.
The encoding makes use of three OOP abstraction mechanisms:
\emph{inheritance}, \emph{subtyping} and \emph{type-refinement}.
Specifically, the new \lstinline{Circuit} is a subtype of
\lstinline{base.Circuit} and declares a new method \lstinline{wellSized}.
The hierarchy implements the new \lstinline{Circuit} by inheriting the corresponding class
from \lstinline{base} and
complementing the body of \lstinline{wellSized}.
Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit}
type to avoid type mismatch in creating objects.

>>>>

@@

Conversely, we can introduce new interpretation in a
modular and intuitive way with an OO language like Scala:
\begin{figure}
\lstinputlisting[linerange=2-15]{../src/wellsized/Circuit.scala}%APPLY:MULTIPLE_SCALA
\caption{Adding new interpretations}
\label{code:operation}
\end{figure}
Instead of modifying the original code, we define \lstinline{wellSized} modularly.
The encoding makes use of three OOP abstraction mechanisms:
\emph{inheritance}, \emph{subtyping} and \emph{type-refinement}.
Specifically, the new \lstinline{Circuit} is a subtype of
\lstinline{base.Circuit} and declares a new method \lstinline{wellSized}.
The hierarchy implements the new \lstinline{Circuit} by inheriting the corresponding class
from \lstinline{base} and
complementing the body of \lstinline{wellSized}.
Also, fields of \lstinline{Beside} are refined with the new \lstinline{Circuit}
type to avoid type mismatch in creating objects.

### 2016-06-21 paper/sections/Transformations.tex
Clean up code
Shorten S4

\section{Transformations in Shallow Embeddings}
Transformations are typically regarded as the privilege of deep embedding.
Since there is not a data structure persisting the AST, how to define transformations
becomes unclear, which becomes one of the main criticism of shallow embedding.
It is not completely true especially for the OO setting.
This section illustrates how to define various transformations in shallow embedding.
%In fact, transformations can still be defined in shallow embedding.

\subsection{Desugaring}
Desugaring is one kind of transformations, which eliminates some
language constructs that could be expressed by other constructs without
sacrificing the expressiveness of the whole language.
In \dsl, the \emph{identity n} construct is such a syntactic sugar that can
be rewritten via the following formula:
$$identity\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
which states that \emph{identity n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
%Using the above formula, we can define a transformation that desugars every
%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.

Figure~\ref{code:desugar} shows how to implement desugaring in Scala.
\begin{figure}
\lstinputlisting[linerange=4-25]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
\caption{OO implementation of desugaring}
\label{code:desugar}
\end{figure}
The newly introduced method \lstinline{desugar} in \lstinline{Circuit} returns a \lstinline{Circuit} reflecting that it is a transformation. The implementation of \lstinline{desugar} for each case is
straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
their results into an instance of \lstinline{Beside}; \emph{identity n} mimics the
formula.

@@

\subsection{Sophisticated Optimizations}
Sophisticated optimizations may need to inspect multiple internal
representations of values.
Suppose we want to merge consecutive vertical wires together:
$$
beside\ (identity\ m)\ (identity\ n) = identity\ (m + n)
$$
This rewrite rule applies only when both $c_1$ and $c_2$ of the $beside\ c_1\ c_2$
are $identity$, which has to inspect the internal representation of $c_1$ and $c_2$ simultaneously.
However, pattern matching can only be simulated only at the the top level, i.e. $beside$.
There is no way to tell what the class types of $c_1$ and $c_2$ are.
Fortunately, Scala has built-in support for pattern matching. Nevertheless, we
can not simply decorate classes with \lstinline{case} modifier as
it is not possible to have a case class that extends another case class.
Alternatively, we define the \lstinline{unapply} method, a.k.a \emph{extractor}, in the companion object
to tell the compiler how to destruct the object. Here comes the implementation
of the above optimization:
\lstinputlisting[linerange=4-22]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA

>>>>

\section{Transformations in Shallow Embeddings}
Transformations are typically regarded as the privilege of deep embeddings.
Since there is not a data structure persisting the AST, how to define transformations
becomes unclear.
It is not completely true, especially for the OO approach.
This section illustrates how to define various transformations, the missing
kind of interpretations in Gibbons and Wu's paper, in shallow embeddings.
%In fact, transformations can still be defined in shallow embedding.

\subsection{Desugaring}
Desugaring is one kind of transformations, which eliminates some
language constructs without sacrificing the expressiveness of the whole language.
In \dsl, the \emph{id} construct is such a syntactic sugar that can
be rewritten via the following formula:
$$id\ n = \overbrace{\ beside\ (fan\ 1)}^{\text{repeat }n-1\text{ times}}\ (fan\ 1)$$
which states that \emph{id n} can be represented as \emph{n} of \emph{fan 1} combined with \emph{beside}.
%Using the above formula, we can define a transformation that desugars every
%occurrence of \emph{identity} to a combination of \emph{fan} and \emph{beside}.

Figure~\ref{code:desugar} shows how to eliminate \lstinline{Id} construct in Scala.
\begin{figure}
\lstinputlisting[linerange=4-23]{../src/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
\caption{OO implementation of desugaring}
\label{code:desugar}
\end{figure}
The newly introduced method \lstinline{desugar} in \lstinline{Circuit} returns a
\lstinline{Circuit} reflecting the nature of a transformation. The implementation of \lstinline{desugar} for each case is
straightforward: \emph{fan n} returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
their results into an instance of \lstinline{Beside}; \emph{id n} just mimics the
formula.

@@

\subsection{Sophisticated Optimizations}
Sophisticated optimizations may need to inspect multiple internal
representations of values.
Suppose we want to merge consecutive vertical wires:
$$
beside\ (id\ m)\ (id\ n) = id\ (m + n)
$$
This rewrite rule applies only when both $c_1$ and $c_2$ of the $beside\ c_1\ c_2$
are $id$, which has to inspect the internal representation of $c_1$ and $c_2$ simultaneously.
However, pattern matching can only be simulated only at the the top level, i.e. $beside$.
There is no easy way to destruct $c_1$ and $c_2$.
Fortunately, Scala has built-in support for pattern matching. For extensibility reasons, we
can not simply decorate classes with \lstinline{case} modifier.
% as it is not possible to have a case class that extends another case class.
Alternatively, we manually define the \lstinline{unapply} method, a.k.a \emph{extractor}, in the companion object. Here comes the implementation
of the above optimization:
\lstinputlisting[linerange=4-22]{../src/optimizations/Circuit.scala}%APPLY:MERGEIDS_SCALA
